To Airbnb?: A Question of Returns

### 2018-05-26 papers/full_paper.Rmd

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one.

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental market in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data.

@@

Existing research into short-term rental markets is sparse. However, with the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.  Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus resulted from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast, the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability toward owners.  The authors also note that very little research has been put toward understanding this growing market segment.

@@

A critical theme that arises out of the extant literature is that fundamentals of the markets for primary dwellings (owner-occupied and long-term rentals) may differ from those of vacation homes (or, by association, short-term rentals). Or, in other words, the factors that influence prices and rents in the two are not equivalent. @cho2003impacts find that the second home market is more sensitive to proximity to environmental amenities than the owner occupied market. In Switzerland, @soguel2008impact test to see if second homes pay a higher premium for ski resort access.  Their results fail to show a statistically significant difference in the impact of ski resorts between markets, but do find that other property characteristics have heterogenous effects between the two markets. In examining the difference between the urban and rural STR markets,  @nilsson2015influence makes a key distinction finding that rural vacation home markets value amenities whereas urban holiday properties tend to derive large premiums from proximity to attractive residential areas.

Though issues of second homes have been primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism'' Fuller and Michel [@fuller2014stop] has re-focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary though Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, impacts on housing market and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy. The closest analog in the literature is a piece by @larson2009pur examining the choice between purchasing a time-share or simply paying for hotel rooms.  Theoretically, the literature shows that in most cases there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that this same thinking can be extended to the difference between short-term rental occupants (tourists and business travelers) and long-term rental occupants (workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term leasing strategy.

@@

We approach answering the question of whether it is more profitable to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a property.  More specifically, we examine the following case:

@@

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we attempted to also estimate short-term rental revenue for long-term properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item and directly paid by the occupants they can also be ignored in the comparative analysis below as they are direct.

@@

We will consider short-term rental (STR) leasting strategy to be prefable to long-term rental (LTR) strategy if the following inequality holds:

@@

Though the analysis we will refer to any situations where the short-term net revenue exceeds the long-term net revenue as 'Short-Term Preferable' or a situation of 'Short-Term Preference'.  For the purposes of this study we consider this a binary position leading to a binary decision of which leasing strategy to follow.

@@

Data for this study were gathered from two different sources, one for short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas as well and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites. The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property-level and nightly-per-property. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period.^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  The nightly data indicate the advertised price, the occupancy status and a unique reservation identification.

@@

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include address. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on a two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long-term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental (LTR) versus an Airbnb-type, short-term rental (STR) approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the short-term data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units or Studios, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).^[Also note that 'Apartment-type' ownership is very rare in Australia.  Nearly every multi-family dwelling is owned as a condominium.  See @berry2000investment for more information on the overall structure of the Australian market] Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, likely, Townhomes in the North American context.

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two. From this analysis we see that there are 6 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2, 3/3, 4/2 and 4/3.  The four bedrooms options are only present in the house category. We filter the data to these combinations.

@@

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10 to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts. The specific suburb of a property is the second, or finer scale, at which the market operates.

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (inner, middle and outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. Rural (Outer suburbs, not Beach)
2. Suburban (Middle suburbs, not Beach)
3. City (Inner Suburbs, not Beach, not Core)
4. City-Core (Select inner suburbs with tourist activities and near CBD)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we also remove any observations that fall outside of outer suburbs (the extent of the suburbs). The location of the short- and long-term rentals, colored by submarket, are shown in .

@@

 To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (\autoref{smcounts}). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

Within the remainder of this report when we refer to revenues we mean the net revenues from equations 1 and 3, respectively. Calculations of revenue for the observed and hypothetical long-term rentals are straightforward, weekly rent -- real or imputed -- times 52 minus search cost time and agency fees.  For the short-term rentals it is somewhat more complicated.

@@

1. Actual ($REV_o$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long.^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better measure of actual market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties. Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year (\autoref{revdens}).  Actual short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Actual measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.

@@

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type, bed/bath and geographic area. In other words, we separately estimate LTR rents for Apartments with 1bed/1bath in the city core, Apartments with 1bed/1bath in the beach submarket, etc. for all combinations of type, bed/bath and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.

@@

The analyses above show that, overall, more than one out of every four short-term rental properties was actually more profitable as a short-term rental than it would have been if rented long-term over the course of the year. This proportion, however, varied by location and property type, with apartments and city-core properties showing the highest likelihood of short-term preference. While illustrative, the above do not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

We specify three, increasingly complex, logistic regression models. We use the actual revenue (not potential) as the dependent variable here, saving an anaylsis of potential revenue for the subsequent sensitivity tests.

@@

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. There is very little significant difference between 1 bed / 1 bath units and the remaining categories, with the exception of 2 bed / 2 bath. This could be due to the fact that 2 bed / 2 bath units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Additionally, short-term tenants may not appreciate (pay) for the second bathroom as extra bathrooms are less of amenity during a short vacation stay than in day-to-day living.

@@

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a weak, but positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sets of sensitivity tests.  First, we examine how the results change if Potential revenue, not Actual revenue, of the short-term properties is used in the comparisons.  Second, we categorize the hosts, split the data and re-analyze the performance of properties based on hypothetical motivations and actions of the hosts.

@@

The above uses the Actual^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to not factoring in the impact of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units in the market that would occur if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 53.3% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 5) shows similar findings, in terms of relative comparativeness, to the actual revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

Our final test of the potential revenue case is to re-run the logistic regression models using short-term preference in the potential case as the dependent variable. Generally speaking, the results from the new series of models (Table 6) are qualitatively similar to those derived in the Actual revenue case (Table 4).  The overall magnitude of the coefficient for House (vs Apartments) is lowered in the Potential case, suggesting that houses may be blocked off more that Apartments.  The product types (bed/bath) show similar trends, though with larger discounts for the 2 Bed / 1 Bath and 2 Bed/2 Bath units.  Likewise the geographic submarkets and host policies show similar effects in terms of direction of impact with some slight deviations in magnitude.

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Actual case and a much smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts are already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences for host types in the Actual number going forward.

@@

#### Data, Code and Interactivity

>>>>

With the rise of Airbnb and other self-hosting accommodation platforms^[Flipkey.com, VRBO.com, Homeaway.com, Tripping.com, Stayz.com.au as examples.], it is easier than ever for property owners and investors to consider short-term rental (STR) as a potential leasing strategy for a revenue generating residential property. Should they though? Or rather, is it more profitable to do so when compared to traditional long-term rental (LTR)? And further, are certain types of properties in specific locations more likely to show a preference for a short-term leasing strategy over long-term one.

To address these questions and to provide some guidance to potential and current investors and owners, we examine the short- and long-term residential rental markets in Melbourne, Australia. Melbourne has a thriving short-term rental market with over 20,000 total properties listed on the Airbnb portal, the source of our short-term market data.

@@

Existing research into short-term rental markets is sparse. Within the existing work, the focus remains centered on rural and/or amenity-based STRs such as beach-front and ski-resort access properties.  This focus resulted from the fact that, until a few years ago, the STR market itself operated almost wholly within these market segments. The shift to small, urban dwellings within the STR market is relatively new and explains the dearth of past work in this direction. With the arrival of Airbnb, and the controversies surrounding it, the topic is creating interest both in real estate and economics as well as in a wide variety of fields as evidenced by a growth in recent publications and on-going working papers.

@@

Focusing specifically on the short-term rental market, @cassidy2007tourist suggest that within Australia's Gold Coast, the STR industry lacks sophistication and efficiency.  Most notably, they find that most nightly or weekly rates are set by resident unit managers (RUMs) and are often under-priced as there is little accountability to owners.  The authors also note that very little research has been put toward understanding this growing market segment.

@@

A critical theme that arises out of the extant literature is that market fundamentals for primary dwellings (owner-occupied and long-term rentals) may differ from those of vacation homes (or, by association, short-term rentals). Or, in other words, the factors that influence prices and rents in the two are not equivalent. @cho2003impacts find that the second home market is more sensitive to proximity to environmental amenities than the owner occupied market. In Switzerland, @soguel2008impact test to see if second homes pay a higher premium for ski resort access.  Their results fail to show a statistically significant difference in the impact of ski resorts between markets, but do find that other property characteristics have heterogenous effects between the two markets. In examining the difference between the urban and rural STR markets,  @nilsson2015influence makes a key distinction finding that rural vacation home markets value amenities whereas urban holiday properties tend to derive large premiums from proximity to attractive residential areas.

Though issues of second homes have been primarily treated as a rural phenomenom for decades  [@gallent2007second], a wave of ``new urban tourism.'' Fuller and Michel [-@fuller2014stop] has re-focused some of the problems of short-term rentals on large, urban environments. With this move to urban areas, issues such as nuisance, property rights and regulation have naturally arisen.  As a result, the short-term rental market has garnered significant interest in the public policy and legal sphere as of late. As many second homes are not permitted for or regulated as commercial entities, STRs can have serious effects on immediate neighbors in the way of externalies such as noise and congestion [@frost2006commentary].  While strict regulation is often proposed, doing so may constititute regulatory takings [@jefferson2015can].  Other, rights-based, regulatory approaches have been suggested [@miller2014transferable].  Many of the key legal decisions regarding short-term rentals, particulary through Airbnb, are still in progress and much remains to be settled in this area in terms of the regulatory and legal environment in which STRs are allowed to operate.

Overall, the extant literature on short-term rentals is limited, but growing.  Of the three types of research -- pricing, impacts on housing market and legalities/externalities -- none specifically address the decision faced by an owner who is trying to decide between a long- or short-term leasing strategy. The closest analog in the literature is a piece by @larson2009pur examining the choice between purchasing a time-share or simply paying for hotel rooms.  Theoretically, the literature shows that in most cases there is a demonstratable difference between first and second home market dynamics due to the use differences of the occupants.  We hypothesize that this same thinking can be extended to the difference between short-term rental occupants (tourists and business travelers) and long-term rental occupants (local workers and retirees) and, as a result, there should be differences in the market which make certain housing products and locations preferable (more profitable) under a short-term versus long-term leasing strategy.

@@

We approach answering the question of whether it is more profitable to pursue a short- or long-term leasing strategy by creating and assessing a scenario likely to be faced by a recent purchaser of a resdiential investment property.  More specifically, we examine the following case:

@@

We do not directly observe properties that are utilized in both a short- and long-term leasing strategy. Therefore, comparisons are done based on estimated or imputed revenues. Most specifically, we will estimate the long-term revenue for observations of properties currently listed on a short-term rental portal (Airbnb). Initially, we attempted to also estimate short-term rental revenue for long-term rental properties but found the process of imputing occupancy rates and nightly rate structures too inaccurate to provide valid comparisons.

@@

Janitorial costs, $J$, cover the costs to 'turn-over' the property between occupants.  For a short-term leasing strategy, these costs can be experienced every day or two, depending on the minimum length of stay and the total number of occupancies over the course of the year. Our data for short-term rentals comes from the Airbnb platform.  In addition to the nightly rate that an occupant pays to the owner there is also a separate 'cleaning fee' that is attached, as is a booking fee to Airbnb.^[This fee could be considered analogous to the leasing fees $L$ that long-term rentals pay.] Some hosts add additional 'linen' fees to cover costs of laundering.  While it may be advantageous for hosts in competitive markets to underquote their cleaning fees to keep occupancies up, we have no evidence to suggest that owners are exhibiting such behavior. As a result of janitorial fees being a separate line item in the STR data and directly paid by the occupants they can also be ignored in the comparative analysis below.

@@

We will consider a short-term rental (STR) leasing strategy to be preferable to a long-term rental (LTR) strategy if the following inequality holds:

@@

Throughout this analysis we will refer to any situations where the short-term net revenue exceeds the long-term net revenue as 'Short-Term Preferable' or a situation of 'Short-Term Preference'.  For the purposes of this study we consider this a binary position leading to a binary decision of which leasing strategy to follow.

@@

Data for this study were gathered from two different sources, one for the short-term rental market and one for the long-term rental market.  Short-term rental market observations are from the Airbnb online portal. Airbnb is the largest short-term rental hub in the Melbourne region. As opposed to similar sites such as VRBO.com and Homeaway.com which focus on vacation homes in tourist dominated areas such as beach towns and ski resorts, Airbnb offers considerable coverage in urban areas and, therefore, directly competes to a greater degree with more traditional long-term rentals than the other sites. The data on the Airbnb market in Melbourne have been purchased from www.airdna.co, a data provider specializing in Airbnb data collection and analysis. The short-term data include two types of observations, property-level and nightly-per-property. Property-level data consists of a single observation for each property that is or has been listed on the Airbnb portal over the October 2014 to August 2016 period.^[This time period represents the period in which AirDNA was engaged in data collection in Melbourne.]  The nightly data offer one observation for each property for each night of the year during its time on the portal.  The nightly data indicate the advertised price, the occupancy status and a unique reservation identification.

@@

Unfortunately, there are no common identifiers between the two datasets as the Airbnb data does not include address. Both sets of data do include latitude and longitude, but in a dense urban area with many multiple-family dwellings it is impossible to match apartments based on two-dimensional coordinates. As a result of this, we are unable to directly match observations from the two datasets. However, the number of properties that rented long-term for the 2014 to 2015 period and then switched to Airbnb for the 2015 to 2016 period are likely very small and this inability to directly match properties does not greatly impact our modeling strategy.

@@

We begin by filtering observations by time. Our hypothetical leasing strategy decision is that of an investor who purchases a property on September 1, 2015 and must decide on traditional long-term rental versus an Airbnb-type, short-term rental approach to generating income. To determine the likely LTR rate for any given property, we will use the observed rental transactions from September 1, 2014 to August 31, 2015. Taking the Airbnb or STR approach allows the owner to change rates over time and is subjected to daily changes in the market (supply and demand factors). In this simulated example we will use Airbnb data from September 1, 2015 to August 31, 2016 to represent the actual Airbnb market conditions over the time period in question.

@@

Part of the difficulty in perfectly mapping property types from these two datasets is due to the fact that the long-term data uses Australian terms while the short-term data conforms to North American lexicon.  This is most apparent when talking about apartment dwellings.  In Australia these are referred to as Units or Studios, while in North American they are referred to as Apartments or Condominiums (depending on ownership structure).^[Also note that 'Apartment-type' ownership is very rare in Australia.  Nearly every multi-family dwelling is owned as a condominium.  See @berry2000investment for more information on the overall structure of the Australian market] Additionally, Terrace homes are very common in many of the older suburbs of Melbourne and would be considered Rowhouses or, possibly, Townhomes in the North American context.

@@

In addition to looking at bedrooms and bathrooms as separate dimensions of properties, they can also be considered together.  In other words, it is very common in the industry (and when looking for a short-term rental) to specify searches by bed/bath combination for obvious reasons -- 1 bedroom with 3 baths is inefficient and 4 bedrooms with 1 bath is uncomfortable.  As a result we examine the combinations of the two. From this analysis we see that there are 8 combinations that make up most all properties:  1bed/1bath, 2/1, 2/2, 3/1, 3/2, 3/3, 4/2 and 4/3.  The four bedrooms options are only present in the house category. We filter the data to these combinations.

@@

Location is an important determinant in both the short- and long-term rental markets. At the broadest scale, the long-term Melbourne residential market is usually discussed in terms of Inner, Middle and Outer suburbs. Inner suburbs are < 10km from the CBD, Middle suburbs 10km to 20km from CBD and Outer suburbs > 20km from CBD. In general, prices are highest in the inner suburbs and lowest in the outer, with a number of exceptions in the high-end neighborhoods in the east and southeastern areas of the middle suburbs.  Suburbs in Melbourne are much smaller than their North American counterparts.

It is likely that the short-term rental market follows a similar market hierarchy, however, we can imagine that a few different spatial features can influence the short-term market.  While there are many possible additional features, we consider two to be the most likely to influence rates or occupancy: 1) Proximity to beaches; and 2) Proximity to key tourist activities and events. Using the three broad submarkets from the long-term market (Inner, Middle and Outer) we have created a five sub-market system as a starting point for spatially analyzing short and long-term rentals in Melbourne:

1. City-Core (Select inner suburbs with tourist activities and near CBD)
2. City (Inner Suburbs, not Beach, not Core)
3. Suburban (Middle suburbs, not Beach)
4. Rural (Outer suburbs, not Beach)
5. Beach (Properties within 500m of Port Phillip Bay east of Yarra River)

@@

We assign these designations by: 1) Adding suburb designations to the properties; 2) Assigning submarkets 1-4 based on suburb location; and 3) Indicating proximity to beach and labeling as 'Beach' submarket.  This process is repeated for both the short- and long-term data. Finally, we filter our data to observations in suburbs that contain at least three STR and LTR observations. The location of the short- and long-term rentals, colored by submarket, are shown in Figures 2 and 3.

@@

To better visualize the relative frequency of the submarkets in the two tenure types we have broken down the data by submarket by property type by tenure type. A number of interesting observations are illustrated here (Figure 4). First, the vast majority of Airbnb units are apartments in city-core or city locations. As expected and as shown by the map, the long-term properties are more evenly spread across the metro region than the short-term properties.  Long-term houses are more peripheral than apartments, a result of urban development patterns and land economics. Beach properties are, relatively speaking, over-represented in the short-term market, especially in apartments, while suburban and rural apartments, somewhat common in long-term markets, are nearly absent in the short-term market. Overall, the suburban and rural Airbnb markets are quite thin, later results focusing on these markets will need to be examined in light of this small sample size.

@@

Within the remainder of this study when we refer to revenues we mean the net revenues from equations 1 and 3, respectively. Calculations of revenue for the observed and hypothetical long-term rentals are straightforward, weekly rent -- real or imputed -- times 52 minus search cost time and agency fees.  For the short-term rentals it is somewhat more complicated.

@@

1. Actual ($REV_o$)}: Observed bookings multiplied by median nightly rate
2. Likely ($REV_l$)}: Observed bookings extrapolated to a full year representing the likely actual revenue if the property had been listed all year long,^[Properties listed in summer, but not winter may be over imputed and vice versa, however, we expect the biases to cancel out over the entire sample.] based on observed occupancy rate multiplied by median nightly rate
3. Potential ($REV_p$)}: Observed occupancy rate of non-blocked days multiplied by 366 multiplied by median nightly rate

As the 'Actual' calculation systematically under counts the revenue from properties that were not on the Airbnb portal for the entire year we disregard it going forward and consider the 'Likely' revenue metric to be a better measure of actual market performance. We chart the distribution of the Likely ($REV_l$) and Potential ($REV_p$) short-term revenues against a distribution of the observed long-term revenues for the long-term properties (Figure 5). Revenues from the long-term rentals are far more consistent, with the vast majority totaling between \$12,000 to \$30,000 per year.  Likely short-term revenues range from neglible to highs of greater than \$80,000, with many of the properties showing very limited (less than \$10,000) revenue. The range of the distribution for the Potential short-term revenue is similar, however, the distribution itself is much flatter, with many more properties showing potential incomes in the \$20,000 to \$40,000 range.

@@

In this section we compare net revenues between long- and short-term rental options.  As mentioned above, there are no direct matches between the datasets. In other words, we have not directly observed a property which has been in a long-term rental followed by short-term rental or vice versa.  We will, instead, make the comparisons between the two by imputing likely long-term rental revenues for the observed short-term properties and then comparing the measured net revenues of the short-term properties to their estimated imputed long-term revenues. We first focus on using the Likely measure of net revenue for the short-term properties as discussed above.  We save discussion and analysis of the Potential revenue for the sensitivity tests later in the study.

@@

Realizing that the imputed long-term rental values may be more accurate when estimated at smaller scales, we impute long-term rental rates at submarket level, specifically by type and geographic area. In other words, we separately estimate the LTR model in Equation 10 for Apartments in the city core, Apartments with  in the beach submarket, etc. for all combinations of type and geographic area. We then combine the results from all estimates back into a single dataset that contains all of the short-term rentals.

@@

Overall, 28.4% short-term rental properties were more profitable as a short-term rental than it would have been if rented long-term over the course of the year. While illustrative, a simple univariate analysis does not fully control for property characteristics or owner actions -- which are a key component of short-term rental success.  To better understand which factors are related to a higher likelihood of short-term preference, we develop a set of logistic regression models below.

We specify three, increasingly complex, logistic regression models. We use the likely revenue (not potential) as the dependent variable here, saving an anaylsis of potential revenue for the subsequent sensitivity tests.

@@

In Model 1, we see that when controlling for structural characteristics, Houses have a significantly lower probability of being more profitable in short-term leasing than apartments. Additionally, 1 bed / 1 bath units show greater likelihood of short-term preference than 2 bed / 1 bath and 2 bed / 2 bath combinations. This could be due to the fact that 2 bed units may be very attractive in the long-term market as they are a useful product for two, non-family roomates to share. Larger units, 3 bedrooms plus, generally show no statistical difference in preference porbability to the 1 bed / 1 bath reference unit, with the exception of the 4 bed / 2 bath properties.

@@

Finally, in Model 3 we add in host behavior or policies. Allowing a greater number of guests per bedroom results in a higher probability of short-term preference.  This variable may be operating as proxy for additional sleeping spaces which allow for a higher nightly rate, ceteris paribus, to be charged.  The greater the minimum stay, the lower the probability of short-term preference.  Longer minimum stays may equate to lower occupancy rates as one or two night travelers are exempted from the property.  Owners with longer minimum stays may be looking to minimize costs such as cleaning and administration.  Flexible cancellation policies result in a lower probability of short-term preference, while a Strict policy has a positive relationship with short-term preference.  Strict policies do help maintain revenues, but may also scare off some potential guests.  Strict cancellation may be a luxury for the best and highly desired properties on the market and, within our model, may be acting as a proxy for some unobserved measure of quality or demand while a flexible cancellation may be signalling the opposite.

@@

To test if the above results are influence by the assumptions we have made throughout this analysis, we perform two sets of sensitivity tests.  First, we examine how the results change if Potential revenue, not Likely revenue, of the short-term properties is used in the comparisons.  Second, we categorize the hosts, split the data and re-analyze the performance of properties based on hypothesized motivations and actions of the hosts.

@@

The above uses the Likely^[Extrapolated out in cases of Airbnb properties on the market for less than a year] revenue in the comparisons with the imputed long-term revenue.  Doing so may underestimate the possible revenue -- what may matter most to the hypothetical buyer in our scenario -- due to not factoring in the impact of blocked days.  If we, instead, assume that blocked days could be leased out on the short-term market at the same overall occupancy rate as the unblocked (observed) days we can produce a potential revenue figure. Note that it is possible (likely?) that the large oversupply of available units in the market that would occur if no properties blocked their units would decrease both prices and occupancy rates.  We do not attempt to correct for this likelihood, rather we offer these figures as high end estimates.

Using the potential revenue, 53.3% of all properties are short-term preferable.  Breaking that down by type and by geographic submarket (Table 3) shows similar findings, in terms of relative comparativeness, to the Likely revenue analysis above.  Apartments have a higher short-term preference than do Houses and City-Core properties dominate the other geographic regions.  One notable difference here is that the Rural properties perform rather poorly under a potential revenue scenario, relatively speaking, suggesting that they may be operating near their peak in the current condition.

@@

Next, we re-run the logistic regression models using short-term preference in the potential case as the dependent variable. Generally speaking, the results from the new series of models (Table 4) are qualitatively similar to those derived in the Likely revenue case (Table 2).  The overall magnitude of the coefficient for House (vs Apartments) is lowered in the Potential case, suggesting that houses may be blocked off more that Apartments.  The product types (bed/bath) show similar trends, though with larger discounts for the 2 Bed / 1 Bath and 2 Bed/2 Bath units.  Likewise the geographic submarkets and host policies show similar effects in terms of direction of impact with some slight deviations in magnitude.

@@

Looking at the differences in short-term preference under both revenue cases, we see large variation in short-term preference between host types in the Likely case and a much smaller variation in the Potential case, as expected. There is little change for the 'Profit Seekers' as most of these hosts are already operating with few blocked days.  For the other three sets, considerable gains could be had with a more aggresive leasing approach. As the potential figures are somewhat similar here, we will only examine the differences for host types in the Likely number going forward.

@@

#### Data and Code
